The hedging hyperprior for the normal distribution

### 2023-03-07 text/23/normal-hedging-hyperprior.Rmd
Started work on normal-hedging-hyperprior

---
title: "The hedging hyperprior for the exponential distribution"
source: "New"
date: "2023-02-28"
categories:
- Blog post
tags:
- Bayesian statistics
output: html_document
---

@@

I want to write a paper about the hedging hyperprior. I need to work out some simple examples first. Here is an example involving the exponential distribution.

@@

Consider a Bayesian analysis of an outcome variable that has an exponential distribution. The density of the exponential distribution is

$f(x) = \theta^{-1} e^{-x / \theta}$

You have an informative prior on $\theta$ that has an inverse gamma distribution

$g(\theta; \alpha, \beta)=\frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{-\alpha-1}e^{-\beta / \theta}$

@@

$L(\mathbf{X})=\prod_{i=1}^n f(X_i)=\theta^{-n} e^{-n \bar X / \theta}$

and a posterior distribution of

$h(\theta | \mathbf{X}) \propto \frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{-\alpha-1}e^{-\beta / \theta}\theta^{-n} e^{-n \bar X / \theta}$

@@

$h(\theta | \mathbf{X}) \propto \theta^{-\alpha-1-n}e^{-(\beta + n\bar X) / \theta}$

>>>>

---
title: "The hedging hyperprior for the normal distribution"
source: "New"
date: "2023-02-28"
categories:
- Blog post
tags:
- Bayesian statistics
output: html_document
---

@@

To do:

+ Rewrite normals using a precision parameter
+ Change the hyperparameter from lambda to delta, kappa, tau, psi, or omega.

I want to write a paper about the hedging hyperprior. I need to work out some simple examples first. Here is an example involving the normal distribution.

@@

There are lots of guides on the Internet on using a normal prior for data that is normally distributed with the population standard deviation $\sigma$ known or unknown. I used papers written by [Kevin P. Murphy][mur1] and [Michael I. Jordan][jor1] to get these formulas.

[jor1]: https://people.eecs.berkeley.edu/~jordan/courses/260-spring10/lectures/lecture5.pdf
[mur1]: https://www.cs.ubc.ca/~murphyk/Papers/bayesGauss.pdf

For simplicity, all sums and products are assumed to go from i=1 to n.

Consider a Bayesian analysis of an outcome variable that has a normal distribution. The density of the normal distribution is

$f(X) = \frac{1}{\sigma \sqrt{2 \pi}}e^{-(X-\mu)^2 / 2\sigma^2}$

Assume for simplicity that $\sigma$ is known. A more general case, with $\sigma$ unknown is described on a separate web page.

You have an informative prior on $\mu$ that also has a normal distribution

$g(\mu; \mu_0, \sigma_0)=\frac{1}{\sigma_0 \sqrt{2 \pi}}e^{-(\mu-\mu_0)^2 / 2\sigma_0^2}$

Simplify this by removing any constants not involving $\mu$.

$g(\mu; \mu_0, \sigma_0) \propto e^{-(\mu-\mu_0)^2 / 2\sigma_0^2}$

@@

$L(\mathbf{X} | \mu) = \frac{1}{(\sigma \sqrt{2 \pi})^n}e^{-\sum(X_i-\mu)^2 / 2\sigma^2}$

Rewrite this by incorporating $\bar{X}$ into the squared term

$L(\mathbf{X} | \mu)= \frac{1}{(\sigma \sqrt{2 \pi})^n}e^{-\sum(X_i-\bar{X}+\bar{X}-\mu)^2 / 2\sigma^2}$

$L(\mathbf{X} | \mu) = \frac{1}{(\sigma \sqrt{2 \pi})^n}e^{-((n-1)S^2 / 2\sigma^2}e^{-\sum(\bar{X}-\mu)^2) / 2\sigma^2}$

Pull out anything that is constant with respect to $X_i$ or $\mu$.

$L(\mathbf{X}| \mu) \propto e^{-\sum(\bar{X}-\mu)^2) / 2\sigma^2}$

The posterior distribution is proportional to the product of the prior and the likelihood.

$h(\mu | \mathbf{X}) \propto e^{-(\mu-\mu_0)^2 / 2\sigma_0^2}e^{-n(\bar{X}-\mu)^2) / 2\sigma^2}$

@@

$h(\mu | \mathbf{X}) \propto e^{-((\frac{\sigma}{\sigma_0}(\mu_0-\mu))^2 + n (\bar{X}-\mu)^2) / 2\sigma^2}$
